---
title: "DATA 608 - Assignment 1"
author: "Rachel Greenlee"
output:
  html_document: default
  html_notebook: default
---

```{r}
library('dplyr')
library('ggplot2')
```


**Principles of Data Visualization and Introduction to ggplot2**

I have provided you with data about the 5,000 fastest growing companies in the US, as compiled by Inc. magazine. lets read this in:

```{r}

inc <- read.csv("https://raw.githubusercontent.com/charleyferrari/CUNY_DATA_608/master/module1/Data/inc5000_data.csv", header= TRUE)

```

And lets preview this data:

```{r}

head(inc)

```



```{r}

summary(inc)

```

Think a bit on what these summaries mean. Use the space below to add some more relevant non-visual exploratory information you think helps you understand this data:

Taking a look at the top 10 and bottom 10 ranked companies.

```{r}
head(inc, 10)
tail(inc, 10)
```
A quick look at the locations of these companies and the frequency in each state. Only 134 in my current state of Colorado, and a mere 79 in my home state of Wisconsin.

```{r}
inc %>% count(State, sort = TRUE)
```

Using similar code we can look at the most common industries on the ranking.

```{r}
inc %>% count(Industry, sort = TRUE)

```




## Question 1

Create a graph that shows the distribution of companies in the dataset by State (ie how many are in each state). There are a lot of States, so consider which axis you should use. This visualization is ultimately going to be consumed on a 'portrait' oriented screen (ie taller than wide), which should further guide your layout choices.

```{r, fig.height=8, fig.width=6}

# create 2-variable df of States & counts
state_freq <- inc %>% count(State, sort = TRUE)

# sort df by count so graph displays ordered and not alphabetically
state_freq$State <- factor(state_freq$State,                                    
                  levels = state_freq$State[order(state_freq$n, decreasing = FALSE)])

# set plot theme for assignment
my_plot_theme <- list(
  theme_classic() +
  theme(plot.background = element_rect(fill = "#EFE7DB"),
        panel.background = element_rect(fill = "#EFE7DB"),
        panel.grid.major.x = element_line(color = "white"),
        axis.title.y = element_text(face = "bold"),
        axis.title.x = element_text(face = "bold")))

# plot generation
ggplot(data = state_freq, aes(x = n, y = State)) +
  geom_col(fill="#FBAA82") +
  labs(title = "California, Texas, and New York have the Most Companies", 
       subtitle = "with California having 181% (314) more companies than second-place Texas",
       x="# of Companies in the Ranking", 
       y="State") +
  scale_x_continuous(limits = c(0, 800), 
                     expand = c(.01, 0.5), 
                     breaks = seq(0, 800, 100)) +
  geom_text(
    aes(x = n, label = n),
    size = 3,
    color = "#4C4C4C", hjust = 0) +
  my_plot_theme

```



## Quesiton 2

Lets dig in on the state with the 3rd most companies in the data set. Imagine you work for the state and are interested in how many people are employed by companies in different industries. Create a plot that shows the average and/or median employment by industry for companies in this state (only use cases with full data, use R's `complete.cases()` function.) In addition to this, your graph should show how variable the ranges are, and you should deal with outliers.

```{r}

# Answer Question 2 here

```



## Question 3

Now imagine you work for an investor and want to see which industries generate the most revenue per employee. Create a chart that makes this information clear. Once again, the distribution per industry should be shown.

```{r}

# Answer Question 3 here

```







